---
title: "Homework 3"
output:
  pdf_document:
    latex_engine: xelatex
header-includes:
  - \usepackage{unicode-math}
date: "2023-10-30"
---
Task 1
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(viridis)
library(lubridate)
library(RColorBrewer)
library(grid)
```

```{r}
# Read the data
vgsales <- read.csv("vgsales.csv")

# Extract games that are available on the PC platform
pc_games <- vgsales %>% filter(Platform == "PC")

# Identify the top 5 publishers with the highest number of PC games
pc_top_5_count <- pc_games %>%
  group_by(Publisher) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count)) %>%
  head(5)

# Print the top 5 publishers
print(pc_top_5_count)
```

```{r}
# Create a summary for the entire data
all_games <- vgsales %>%
  group_by(Genre) %>%
  summarize(Count = n(), .groups = "drop")

# Filter out games from the top 5 publishers
pc_top_5 <- pc_games %>% filter(Publisher %in% pc_top_5_count$Publisher)

# Create histograms
all_plot <- ggplot(all_games, aes(x = Genre, y = Count)) +
  geom_bar(stat = "identity") +
  labs(title = "All Platforms", x = NULL, y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

pc_plot <- ggplot(pc_top_5, aes(x = Genre)) + 
  geom_bar(stat = "count") +
  labs(title = "Top 5 PC Platforms", x = NULL, y = NULL) + theme(element_blank()) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

all_plot <- all_plot + theme(plot.title = element_text(size = 10)) 
pc_plot <- pc_plot + theme(plot.title = element_text(size = 10))

# Arrange plots side by side with a bigger title
grid.arrange(all_plot, pc_plot, ncol = 2, top = textGrob("Genre Distribution - All vs. Top 5 PC", gp = gpar(fontsize = 15)))
```

All Platforms: The count for genres spans from around 1000 to over 3000.
Top 5 PC Platforms: The count for genres lies between 0 and 60 because we did filtering.

In All Platforms the "Action" genre as the most dominant. On the other hand "Action" is in the second place after "Shooter" genre on Top 5 PC Platforms. For the top 5 PC platforms, genres like "Shooter," "Action," "Simulation," and "Strategy" are quite popular. This could suggest that PC gamers have a preference for these genres, possibly because PCs can offer more complex controls and richer graphics. We do not see genres like "Fighting" and "Puzzle" on the top 5 PC platforms compared to all platforms. They seem to be not popular for top 5 PC platforms.

```{r}
# Filter data
top_5_data <- vgsales %>% 
  filter(Publisher %in% pc_top_5_count$Publisher) %>%
  group_by(Year, Publisher) %>%
  summarize(Total_Global_Sales = sum(Global_Sales, na.rm = TRUE), .groups = "drop")

# Convert the Year column to numeric and remove NAs
suppressWarnings({
  top_5_data$Year <- as.numeric(as.character(top_5_data$Year))
})
top_5_data <- top_5_data %>% filter(!is.na(Year))

# Plot the line chart
line_chart <- ggplot(top_5_data, aes(x = Year, y = Total_Global_Sales, color = Publisher)) +
  geom_line(aes(group = Publisher), size = 1) +
  labs(title="Global Sales over Years for Top 5 Publishers", 
       x = NULL, 
       y = NULL) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  theme_minimal() + 
  scale_color_brewer(palette ="Set1", name ="Publisher") +
  theme(legend.position ="top", legend.title=element_blank())

# Print the line chart
print(line_chart)
```

The graph provides a comprehensive overview of the global sales trends for five major game publishers over four decades. It captures the rise and fall in sales, indicating both individual publisher strategies and industry-wide influences that have shaped the gaming market.
For example, Activision and Electronic Arts publishers show significant growth, especially after 2005. And Ubisoft's growth is gradual.

```{r}
# Create a cumulative sum of Total_Global_Sales for each publisher
top_5_data <- top_5_data %>%
  arrange(Year) %>%
  group_by(Publisher) %>%
  mutate(Cumulative_Sales = cumsum(Total_Global_Sales))

# Plot the cumulative line chart 
cumulative_chart <- ggplot(top_5_data, aes(x = Year, y = Cumulative_Sales, color = Publisher)) +
  geom_line(aes(group = Publisher), size = 1) +
  labs(title="Cumulative Global Sales over Years for Top 5 Publishers", 
       x = NULL, 
       y = NULL) +
  scale_x_continuous(breaks = seq(1980, 2020, by=5)) +
  theme_minimal() + 
  scale_color_brewer(palette ="Set1", name ="Publisher") +
  theme(legend.position ="top", legend.title = element_blank())

# Print the cumulative line chart
print(cumulative_chart)
```

Electronic Arts has the highest cumulative sales, approaching 900 by 2020.

Activision follows closely, reaching just above 600 by 2020.

Ubisoft and Take-Two Interactive (Purple Line) have a similar growth trajectory, with sales around 500 and slightly below 400 respectively by 2020.

Sega has the lowest cumulative sales among the group, with a relatively flat growth after the early 2000s.

Task 2
```{r}
# Filter and aggregate data for PC by Genre
pc_genre_counts <- vgsales %>%
  filter(Platform == "PC") %>%
  group_by(Genre) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

pc_genre_counts
# Define a custom color palette with 12 colors
colors <- brewer.pal(n = 12, name = "Paired")

# Create pie chart using the custom color palette
pie_chart <- ggplot(pc_genre_counts, aes(x = "", y = Count, fill = Genre)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = colors) +
  labs(title = "Distribution of PC Games by Genres", fill = "Genre") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0.5),
        panel.grid = element_blank(),
        legend.title = element_blank())

print(pie_chart)

print(dominating_genre_pc <- pc_genre_counts[[1,"Genre"]])
```

```{r}
colors_pc <- c("#a6cee3", "#ff7f00", "#cab2d6", "#6a3d9a", "#b15928")

# Filter vgsales to include only PC games, group by Genre, and count the number of games
top_5_pc_genres <- vgsales %>%
  filter(Platform == "PC") %>%
  group_by(Genre) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count)) %>%
  head(5)

# Create a pie chart for PC
pc_pie_chart <- ggplot(top_5_pc_genres, aes(x = "", y = Count, fill = Genre)) +
  geom_bar(stat = "identity", width = 1) +
  geom_text(aes(label = Count), position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = colors_pc) +
  labs(title = "Top 5 Genres - PC Games") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0.5),
        panel.grid = element_blank(),
        legend.title = element_blank())

# Print the pie chart
print(pc_pie_chart)

# Identify the dominating genre for PC
print(dominating_genre_pc <- top_5_pc_genres[[1,"Genre"]])
```
Top 5 Geners PC Games is a subset with Action, Role-Playing, Shooter, Simulation, Strategy from All PC games. Strategy is Dominant.

```{r}
colors_x360 <- c("#a6cee3", "#33a02c", "#fdbf6f", "#cab2d6", "#ffff99")

# Filter and aggregate data for PS4
top_5_x360_genres <- vgsales %>%
  filter(Platform == "X360") %>%
  group_by(Genre) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count)) %>%
  head(5)

# Create a pie chart for PS4
x360_pie_chart <- ggplot(top_5_x360_genres, aes(x = "", y = Count, fill = Genre)) +
  geom_bar(stat = "identity", width = 1) +
  geom_text(aes(label = Count), position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = colors_x360) +
  labs(title = "Top 5 Genres - X360 Games") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0.5),
        panel.grid = element_blank(),
        legend.title = element_blank())

# Print the pie chart
print(x360_pie_chart)

# Identify the dominating genre for X360
print(dominating_genre_x360 <- top_5_x360_genres[[1,"Genre"]])
```
For X360 Games Action, Sports and Shooter have more proportions. 
Both X360 and PC game distributions highlight the popularity of the 'Action' and 'Shooter' genres. the "Action" and "Shooter" genres have higher proportions for X360 games than in the Top 5 PC games. For example, this can indicate a few things:
Xbox 360 players may have a stronger preference for fast-paced and direct engagement games, typical of the "Action" and "Shooter" genres.
The "Sports" genre is more popular in the X360 distribution but doesn't appear in the Top 5 PC games. This could be because sports games are often more enjoyable on consoles due to their multiplayer and local co-op features. 

```{r}
# Get the top 15 publishers
top_15_publishers <- vgsales %>% 
  filter(Platform == "PC") %>% 
  group_by(Publisher) %>% 
  summarise(Count = n()) %>% 
  arrange(-Count) %>% 
  head(15) %>% 
  pull(Publisher)

# Filter the data to only include those top publishers
filtered_data <- vgsales %>% 
  filter(Publisher %in% top_15_publishers)

# Create a table to count games by publisher and genre
count_table <- filtered_data %>% 
  group_by(Publisher, Genre) %>% 
  summarise(Count = n(), .groups = 'drop') %>% 
  pivot_wider(names_from = Genre, values_from = Count, values_fill = 0)

# Convert wide format back to long format for plotting
count_table_long <- count_table %>% 
  pivot_longer(cols = -Publisher, names_to = "Genre", values_to = "Count")
```

```{r}
# Plot the Heatmap
ggplot(count_table_long, aes(x = Publisher, y = Genre, fill = Count)) + 
  geom_tile() + 
  geom_text(aes(label=Count)) + 
  scale_fill_viridis(direction = -1) +  
  labs(title = "Genres by Top 15 PC Game Publishers", 
       x ="Publisher", 
       y = "Genre", 
       fill = "Number of Games") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0.5),
        panel.grid = element_blank(),
        legend.title = element_blank())
```

A publisher is better if it has published more games, then the publishers with more colors tending towards the warmer side of the palette (representing higher game counts) across multiple genres are considered better. In this heatmap, publishers such as Activision, Electronic Arts, and Ubisoft show a diverse portfolio across multiple genres. In this heatmap, publishers such as Activision, Electronic Arts, and Ubisoft show a diverse portfolio across multiple genres.
Electronic Arts seems to have a significant role across multiple genres, with the "Sports" genre being especially dominant, evidenced by the dark shade of purple count of 561.
We see that "Action" stands out as a popular genre across multiple publishers such as Activision, Electronic Arts, THQ, Ubisoft.
Ubisoft have a strong focus on "Action" games, as indicated by the high count of 193.
THQ have a significant count of 194 in the "Action" genre.
Activision have produced a high number of "Action" games, indicated by the count of 310.

Task 4
```{r}
df <- read.csv("MobileSmartphones.csv")
selected_brands <- df %>% 
                  select(Date, Apple, Samsung, Xiaomi, Oppo)
# Convert Date Column
selected_brands$Date <- ym(selected_brands$Date)

# Melt the Data
melted_data <- selected_brands %>% 
               gather(key = "Brand", value = "MarketShare", -Date)

# Determine the start and end dates
start_date <- min(selected_brands$Date)
end_date <- max(selected_brands$Date)

# Plot
ggplot(melted_data, aes(x = Date, y = MarketShare, color = Brand, group = Brand)) +
  geom_line(size = 3) +
  labs(title ="Top 4 Smartphones with Highest Market Share over time",
       y = NULL, 
       x = NULL,
       color="Smartphone") +
  geom_segment(aes(x = start_date, xend = end_date, y = 18.63, yend = 17.41), linetype = "dashed", color = "orange", arrow = arrow(length=unit(1.35, 'cm'))) +
  geom_text(aes(x = start_date, y = 18.63, label = "18.63%"), size = 3.5, hjust = 0, vjust = 1.5, color = "black") +
    geom_text(aes(x = end_date, y = 17.41, label = "17.41%"), size = 3.5, hjust = 1.7, vjust = 1.5, color = "black") +
  geom_text(aes(x = median(selected_brands$Date), y = 20, label = "Tendency of market shares of Top 4"), size = 3.5, hjust = 0.58,color = "black") +
  theme_minimal() +
  scale_x_date(date_breaks = "3 month", date_labels = "%b %Y") + # Format the x-axis for dates
  scale_y_continuous(labels = function(x) paste0(x, "%")) + # Format the y-axis 
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.minor = element_blank())  # This line removes the minor grid lines
```



