---
title: "Homework 4"
output:
  pdf_document:
    latex_engine: xelatex
header-includes:
  - \usepackage{unicode-math}
always_allow_html: yes
date: "2023-11-15"
---

```{r, message = FALSE, echo=FALSE}
library(ggplot2)
library(forecast)
library(lubridate)
library(dplyr)
library(quantmod)
library(xts)
library(stringr)
library(zoo)
library(rnaturalearth)
library(rnaturalearthdata)
library(viridis)
library(gridExtra)
```

## Problem 1
```{r}
monero <- read.csv("coin_Monero.csv")
nemo <- read.csv("coin_NEM.csv")
polkadot <- read.csv("coin_Polkadot.csv")
```


Part 1

Candlestick plots 

```{r, echo=FALSE}
# Convert 'Date' to Date type and sort
monero$Date <- as.Date(monero$Date, format ="%Y-%m-%d %H:%M:%S")
monero <- monero[order(monero$Date),]

# Aggregate data by month
monthly_data_monero <- monero %>%
  mutate(Month = floor_date(Date, "month")) %>%
  group_by(Month) %>%
  summarise(Open = first(Open),
            High = max(High),
            Low = min(Low),
            Close = last(Close))

# Define the candlestick plot function with annotations
candlestick_plot <- function(data) {
  p <- ggplot(data, aes(x = Month)) +
    geom_linerange(aes(ymin = Low, ymax = High), colour = "black") +  
    # Represents the high-low range
    geom_segment(aes(x = Month, xend = Month, y = Open, yend = Close), 
                 size = 1, 
                 colour = "black") +
    geom_rect(aes(xmin = as.Date(Month) - 9, 
                  xmax = as.Date(Month) + 9, 
                  ymin = pmin(Open, Close), 
                  ymax = pmax(Open, Close), 
                  fill = Open < Close),  
              # Determines color based on up or down day
              colour = "black") +
    scale_fill_manual(values = c("TRUE" = "green", "FALSE" = "red")) +
    labs(title = "Monero - Monthly Candlestick", x = "", y = "Price") +
    theme_minimal() +
    theme(legend.position = "none")
  
  highest <- data[which.max(data$High),]
  lowest <- data[which.min(data$Low),]
  
  # Add annotations for the highest and lowest points
  p + geom_text(aes(x = highest$Month, y = highest$High, 
                    label = paste("High:", highest$High)), 
                vjust = -0.02, hjust = 1.2, colour = "black", size = 4) +
      geom_text(aes(x = lowest$Month, y = lowest$Low, 
                    label = paste("Low:", lowest$Low)), 
                vjust = -3, hjust = 0.1, colour = "black", size = 4) +
      # Add arrows pointing to the high and low points
      geom_segment(aes(x = highest$Month, xend = highest$Month, 
                       y = highest$High, yend = highest$High + 5), 
                   arrow = arrow(length = unit(0.5, "cm")), colour = "black") +
      geom_segment(aes(x = lowest$Month, xend = lowest$Month, 
                       y = lowest$Low, yend = lowest$Low - 50), 
                   arrow = arrow(length = unit(0.5, "cm")), colour = "black")
}
```

```{r}
# Call the candlestick plot function
candlestick_plot(monthly_data_monero)
```

```{r}
# Convert 'Date' to Date type and sort
nemo$Date <- as.Date(nemo$Date, format = "%Y-%m-%d %H:%M:%S") 
nemo <- nemo[order(nemo$Date),]

# Aggregate data by month
monthly_data_nemo <- nemo %>%
mutate(Month = floor_date(Date, "month")) %>% group_by(Month) %>%
summarise(Open = first(Open),
High = max(High), Low = min(Low), Close = last(Close))

# Define the candlestick plot function with annotations
candlestick_plot <- function(data) {
  p <- ggplot(data, aes(x = Month)) +
    geom_linerange(aes(ymin = Low, ymax = High), colour = "black") + 
    # Represents the high-low range
    geom_segment(aes(x = Month, xend = Month, y = Open, yend = Close), 
                 size = 1, 
                 colour = "black") +
    geom_rect(aes(xmin = as.Date(Month) - 9, 
                  xmax = as.Date(Month) + 9, 
                  ymin = pmin(Open, Close), 
                  ymax = pmax(Open, Close), 
                  fill = Open < Close),  
              # Determines color based on up or down day
              colour = "black") +
    scale_fill_manual(values = c("TRUE" = "green", "FALSE" = "red")) +
    labs(title = "Nem - Monthly Candlestick", x = "", y = "Price") +
    theme_minimal() +
    theme(legend.position = "none")
  
  highest <- data[which.max(data$High),]
  lowest <- data[which.min(data$Low),]
  
  # Add annotations for the highest and lowest points
  p + geom_text(aes(x = highest$Month, y = highest$High, 
                    label = paste("High:", highest$High)), 
                vjust = 0.5, hjust = 1.2, colour = "black", size = 4) +
      geom_text(aes(x = lowest$Month, y = lowest$Low, 
                    label = paste("Low:", lowest$Low)), 
                vjust = -2, hjust = 0.3, colour = "black", size = 4) +
      # Add arrows pointing to the high and low points
      geom_segment(aes(x = highest$Month, xend = highest$Month, 
                       y = highest$High, yend = highest$High + 0.2), 
                   arrow = arrow(length = unit(0.5, "cm")), colour = "black") +
      geom_segment(aes(x = lowest$Month, xend = lowest$Month, 
                       y = lowest$Low, yend = lowest$Low - 0.2), 
                   arrow = arrow(length = unit(0.5, "cm")), colour = "black")
}
# Call the candlestick plot function
candlestick_plot(monthly_data_nemo)
```

```{r}
# Convert 'Date' to Date type and sort
polkadot$Date <- as.Date(polkadot$Date, format = "%Y-%m-%d %H:%M:%S")
polkadot <- polkadot[order(polkadot$Date),]

# Aggregate data by month
monthly_data_polkadot <- polkadot %>%
  mutate(Month = floor_date(Date, "month")) %>%
  group_by(Month) %>%
  summarise(Open = first(Open),
            High = max(High),
            Low = min(Low),
            Close = last(Close))

# Define the candlestick plot function with annotations
candlestick_plot <- function(data) {
  p <- ggplot(data, aes(x = Month)) +
    geom_linerange(aes(ymin = Low, ymax = High), colour = "black") +  
    # Represents the high-low range
    geom_segment(aes(x = Month, xend = Month, y = Open, yend = Close), 
                 size = 1, 
                 colour = "black") +
    geom_rect(aes(xmin = as.Date(Month) - 9, 
                  xmax = as.Date(Month) + 9, 
                  ymin = pmin(Open, Close), 
                  ymax = pmax(Open, Close), 
                  fill = Open < Close),  # Determines color based on up or down day
              colour = "black") +
    scale_fill_manual(values = c("TRUE" = "green", "FALSE" = "red")) +
    labs(title = "Polkadot - Monthly Candlestick", x = "", y = "Price") +
    theme_minimal() +
    theme(legend.position = "none")
  
  highest <- data[which.max(data$High),]
  lowest <- data[which.min(data$Low),]
  
  # Add annotations for the highest and lowest points
  p + geom_text(aes(x = highest$Month, y = highest$High, 
                    label = paste("High:", highest$High)), 
                vjust = -3, hjust = 1.2, colour = "black", size = 4) +
      geom_text(aes(x = lowest$Month, y = lowest$Low, 
                    label = paste("Low:", lowest$Low)), 
                vjust = -3, hjust = 0.1, colour = "black", size = 4) +
      # Add arrows pointing to the high and low points
      geom_segment(aes(x = highest$Month, xend = highest$Month, 
                       y = highest$High, yend = highest$High + 5), 
                   arrow = arrow(length = unit(0.5, "cm")), colour = "black") +
      geom_segment(aes(x = lowest$Month, xend = lowest$Month, 
                       y = lowest$Low, yend = lowest$Low - 5), 
                   arrow = arrow(length = unit(0.5, "cm")), colour = "black")
}

candlestick_plot(monthly_data_polkadot)
```

Part 2
```{r}
# Create a time series object
monero_ts <- ts(monero$Close, start = c(year(min(monero$Date)),
                                       month(min(monero$Date))),
                end = c(year(max(monero$Date)), 
                        month(max(monero$Date))),
                frequency = 12)

# Create a time series object
nemo_ts <- ts(nemo$Close, start = c(year(min(nemo$Date)), 
                                   month(min(nemo$Date))),
                  end = c(year(max(nemo$Date)), 
                          month(max(nemo$Date))),
                  frequency = 12)

polkadot_ts <- ts(polkadot$Close, 
                  start = c(year(min(polkadot$Date)), 
                            month(min(polkadot$Date))),
                  end = c(year(max(polkadot$Date)), month(max(polkadot$Date))),
                  frequency = 12)
```

Seasonal plots
```{r}
ggseasonplot(monero_ts, year.labels = F) + 
  ggtitle("Seasonal Plot - Monero Close Prices") +
  xlab("") +
  ylab("") +
  theme(panel.grid.minor = element_blank(), 
        plot.title = element_text(size = 14, hjust = 0.5)) +  
  geom_line(linetype = "solid", size = 0.7) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5),
    axis.text.x = element_text(angle = 45, vjust = 1),
    legend.position = "bottom",
    legend.title = element_blank()
  ) 
```

```{r}
ggseasonplot(nemo_ts, year.labels = F) +
  ggtitle("Seasonal Plot - Nem Close Prices")  +
  xlab("") +
  ylab("") +
  theme(panel.grid.minor = element_blank(), 
        plot.title = element_text(size = 14, hjust = 0.5)) +  
  geom_line(linetype = "solid", size = 0.7) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5),
    axis.text.x = element_text(angle = 45, vjust = 1),
    legend.position = "bottom",
    legend.title = element_blank()
  ) 
```

```{r}
ggseasonplot(polkadot_ts, year.labels = F) +
  ggtitle("Seasonal Plot - Polkadot Close Prices")  +
  xlab("") +
  ylab("") +
  theme(panel.grid.minor = element_blank(), 
        plot.title = element_text(size = 14, hjust = 0.5)) +  
  geom_line(linetype = "solid", size = 0.7) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5),
    axis.text.x = element_text(angle = 45, vjust = 1),
    legend.position = "bottom",
    legend.title = element_blank()
  ) 
```
Subseries plot
```{r}
# Plot the subseries plot of the monero time series object
ggsubseriesplot(monero_ts) + ylab("") + 
  ggtitle("Monero Close Prices Over Time by Month") +
  xlab("") + theme_minimal() +
  theme(plot.title = element_text(size = 14, hjust = 0.5))
```

```{r}
# Plot the subseries plot of the nemo time series object
ggsubseriesplot(nemo_ts) + ylab("") + 
  ggtitle("Nem Close Prices Over Time by Month") +
  xlab("") + theme_minimal() +
  theme(plot.title = element_text(size = 14, hjust = 0.5))
```
```{r}

```


Decomposition plots
```{r}
# Assuming monero_ts is your time series object
decomposed <- stl(monero_ts, s.window = 'periodic')

# Get the range of data without NA values for plotting
time_range <- time(na.omit(decomposed$time.series))

# Plot the decomposed object
autoplot(decomposed) +
  ggtitle("STL Decomposition of Monero Close Prices") +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  ) + 
  xlim(min(time_range), max(time_range)) # Set limits to avoid NA values
```

```{r}
# Assuming nemo_ts is your time series object
decomposed <- stl(nemo_ts, s.window = 'periodic')

# Get the range of data without NA values for plotting
time_range <- time(na.omit(decomposed$time.series))

# Plot the decomposed object
autoplot(decomposed) +
  ggtitle("STL Decomposition of Nem Close Prices") +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  ) + 
  xlim(min(time_range), max(time_range)) # Set limits to avoid NA values
```

ACF plots
```{r}
# Plot the ACF of the monero time series object
ggAcf(monero_ts) + ggtitle("ACF of Monero Close Prices") + theme_minimal() +
  theme(plot.title = element_text(size = 14, hjust = 0.5))
```

```{r}
# Plot the ACF of the nemo time series object
ggAcf(nemo_ts) + ggtitle("ACF of Nem Close Prices") + theme_minimal() +
  theme(plot.title = element_text(size = 14, hjust = 0.5))
```

```{r}
ggAcf(polkadot_ts) + ggtitle("ACF of Polkadot Close Prices") + theme_minimal() +
  theme(plot.title = element_text(size = 14, hjust = 0.5))
```
Polkadot contain data less than for 2 years
```{r}
# Plot the line graph for daily close prices of Polkadot
ggplot(polkadot, aes(x = Date, y = Close)) +
  geom_line(linetype = "solid") +
  ggtitle("Polkadot Daily Close Prices") +
  xlab("Date") +
  ylab("Close Price") +
  theme_minimal()+
  theme(plot.title = element_text(size = 14, hjust = 0.5))
```

As a result, decomposition describes the seasoanlity the best because it separates the overall trend, which shows long-term changes, from the recurring patterns that repeat at regular intervals. By isolating these components, decomposition makes it easier to identify and understand the seasonal patterns in the data.
Part 3
```{r}
# Assuming you have defined 'events_monero' with the Monero events data
events_monero <- data.frame(
  Date = as.Date(c('2018-10-19', '2020-10-17', '2017-01-01')),
  Event = c('Bulletproofs Techn', 'Protocol 15', 'Bitfinex')
)

# Create a ggplot for Monero data
p_monero <- ggplot(monero, aes(x = Date, y = Close)) +
  geom_line() +
  theme_minimal() +
  ggtitle("Monero Price with Events")

# Add event annotations for Monero
p_monero <- p_monero +
  geom_segment(data = events_monero, aes(x = Date, xend = Date,
                                         y = Inf, yend = -Inf),
               color = "red") +
  annotate("text", x = events_monero$Date, y = max(monero$Close), 
           label = events_monero$Event, vjust = 1, hjust = 1.5, angle = 90, size = 1.5)


events_nemo <- data.frame(
  Date = as.Date(c('2017-04-01', '2017-09-01', '2021-03-12')),
  Event = c('Nem Technology', 'Chinese Cryptocur. Exchanges', 'Symbol Airdrop')
)

# Create a ggplot for NEMO data
p_nemo <- ggplot(nemo, aes(x = Date, y = Close)) +
  geom_line() +
  theme_minimal() +
  ggtitle("Nem Price with Events")

# Add event annotations for NEMO
p_nemo <- p_nemo +
  geom_segment(data = events_nemo, aes(x = Date, xend = Date,
                                       y = Inf, yend = -Inf),
               color = "red") +
  annotate("text", x = events_nemo$Date, y = max(nemo$Close), 
           label = events_nemo$Event, vjust = 1, hjust = 1, angle = 90, size = 1.5)


events_polkadot <- data.frame(
  Date = as.Date(c('2020-11-01', '2021-05-26')),
  Event = c("Parachain auctions", 'Kusama Network')
)

# Create a ggplot for Polkadot data
p_polkadot <- ggplot(polkadot, aes(x = Date, y = Close)) +
  geom_line() +
  theme_minimal() +
  ggtitle("Polkadot Price with Events")

# Add event annotations for Polkadot
p_polkadot <- p_polkadot +
  geom_segment(data = events_polkadot, aes(x = Date, xend = Date,
                                           y = Inf, yend = -Inf),
               color = "red") +
  annotate("text", x = events_polkadot$Date, y = max(polkadot$Close), 
           label = events_polkadot$Event, vjust = 1, hjust = 1, angle = 90, size = 1.5)

# Combine the three plots vertically
grid.arrange(p_monero, p_nemo, p_polkadot, ncol = 1)

```

Part 4

```{r}
library(ggplot2)
library(gridExtra)

# Monero Close Price Density Plot
monero_plot <- ggplot(monero, aes(x = Close)) +
  geom_density(fill = "red", alpha = 0.5) +
  labs(title = "Monero", x = "", y = "Density") + theme_minimal()

# NEM Close Price Density Plot
nem_plot <- ggplot(nemo, aes(x = Close)) +
  geom_density(fill = "green", alpha = 0.5) +
  labs(title = "Nem", x = "Close Price", y = "") + theme_minimal()

# Polkadot Close Price Density Plot
polkadot_plot <- ggplot(polkadot, aes(x = Close)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Polkadot", x = "", y = "") + theme_minimal()

# Monero Low Price Box Plot
monero_box <- ggplot(monero, aes(x = Low, y = "")) +
  geom_boxplot(fill = "red", alpha = 0.5) +
  labs(title = "", x = "", y = "") + theme_minimal()

# NEM Low Price Box Plot
nem_box <- ggplot(nemo, aes(x = Low, y = "")) +
  geom_boxplot(fill = "green", alpha = 0.5) +
  labs(title = "", x = "Low Price", y = "") + theme_minimal()

# Polkadot Low Price Box Plot
polkadot_box <- ggplot(polkadot, aes(x = Low, y = "")) +
  geom_boxplot(fill = "blue", alpha = 0.5) +
  labs(title = "", x = "", y = "") + theme_minimal()

# Arrange the plots side by side
grid.arrange(monero_plot, nem_plot, polkadot_plot, monero_box, nem_box, polkadot_box, ncol = 3)
```


\newpage

### Problem 2 
```{r datasetup, echo=FALSE}
# Read the CSV file
hotels_data <- read.csv("booking_reviews.csv")

# Count the frequency of each hotel
hotel_counts <- hotels_data %>%
  group_by(hotel_name) %>%
  summarise(Frequency = n()) %>%
  arrange(desc(Frequency)) %>%
  mutate(Index = row_number())

# Filter for the 5th group (j%8 == 4)
group_5 <- hotel_counts %>%
  filter((Index - 1) %% 8 == 4)

# Select the rows from the original dataset for our the hotels 
group_5_hotels <- hotels_data %>%
  filter(hotel_name %in% group_5$hotel_name)

# Convert 'reviewed_at' to Date type
group_5_hotels$reviewed_at <- 
  as.Date(group_5_hotels$reviewed_at, format = "%d %B %Y")
```

```{r}
# Plot the histogram of ratings
ggplot(group_5_hotels, aes(x = rating)) +
  geom_histogram(binwidth = 1, color = "white") +
  labs(title = "Distribution of Ratings",
       x = "",
       y = "Count") + 
  theme_minimal() +  theme(plot.title = element_text(hjust = 0.5))
```
The distribution of all ratings is very left-skewed meaning there are more positive feedback about hotels - the most frequent grades are from 9 to 10. 


\newpage
```{r datasetup2, echo=FALSE}
# Read the countries dataset
countries <- read.csv("countries.csv")

# Select only the 'name' and 'region' columns from the countries dataset
regions <- countries[, c("name", "region")]

# Merge the hotel data with the regions data on the nationality
merged_data <- merge(group_5_hotels, regions, 
                     by.x = "nationality", by.y = "name", 
                     all.x = TRUE)
```



```{r}
# Boxplot of ratings by region
ggplot(na.omit(merged_data[, c("rating", "region")]), 
       aes(x = region, y = rating, fill = region)) +
  geom_boxplot() +
  labs(title = "Relationship Between Regions of Reviewers' Origins and Ratings",
       x = "",
       y = "",
       fill = "Region") +
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```
The boxplot visually summarizes hotel ratings across the regions. They all have median ratings above 7.5, indicating generally positive ratings (it was also shown in the previous plot). Oceania shows the highest median ratings, so we may make a conclusion people with Oceanian origins are more satisfied with the hotels. Median of ratings from African reviewers is lower compared to others suggesting Africans are more strict rating hotels or they may face discrimination based on their race.

\newpage

After testing who is "kinder" when giving reviews to the hotels we are checking people of which nationality gave most reviews.
```{r datasetup3, echo=FALSE}
# Count the number of reviews for each nationality
review_counts <- group_5_hotels %>%
  group_by(nationality) %>%
  summarise(review_count = n())

# Read the countries dataset
countries_data <- read.csv("countries_lat.csv") # Replace with actual path

# Merge review counts with the country data
merged_df <- merge(countries_data, review_counts, by.x = "name", by.y = "nationality")

# Get world map data, excluding Antarctica
world <- ne_countries(scale = "medium", returnclass = "sf") %>%
  filter(name != "Antarctica") 

# Merge world data with review counts
world_with_reviews <- merge(world, merged_df, by.x = "name", by.y = "name", all.x = TRUE)
```

```{r}
# Create the map plot
ggplot() +
  geom_sf(data = world, fill = "white", color = "black") + # Plot all countries in white
  geom_sf(data = world_with_reviews, aes(fill = review_count), color = "black") + # Overlay countries with review counts
  scale_fill_viridis(direction = -1, na.value = NA, option = "D") + # Use reversed viridis palette, with NA for countries without reviews
  theme_void() +
  theme(legend.position = "bottom") + # Move legend to bottom
  labs(fill = "Review Count")
```
We see that most of the reviewers have European origins. Let's zoom in to see in more details from which European countries there are more reviewers.


\newpage
```{r}
xlims <- c(-25, 40) # Adjust these limits for longitude as needed
ylims <- c(34, 72) # Adjust these limits for latitude as needed

# Create the map plot focused on Europe
ggplot() +
  geom_sf(data = world, fill = "white", color = "black") +
  geom_sf(data = world_with_reviews, aes(fill = review_count), color = "black") +
  scale_fill_viridis(direction = -1, na.value = NA, option = "D") +
  theme_void() +
  theme() +
  labs(fill = "Review Count") +
  coord_sf(xlim = xlims, ylim = ylims, expand = FALSE) # Zoom in on Europe
```

There are over 800 Britanian reviewers! Compared to other nationalities it's a lot because most of the countries have about 200-300 reviewers in this dataset. The second most frequent nationality is Belgian. Others are relatively similar.

\newpage
```{r datasetup4, echo=FALSE}
group_5_hotels <- group_5_hotels %>%
  mutate(trip_type = str_extract(tags, "Leisure|Business"))

group_5_hotels$reviewed_at <- 
  as.Date(group_5_hotels$reviewed_at, format = "%d %B %Y")
```

```{r}
ggplot(group_5_hotels[!is.na(group_5_hotels$trip_type), ], 
       aes(x = factor(format(reviewed_at, "%b"), levels = month.abb), 
           fill = trip_type)) +
  geom_bar(position = "stack") +
  labs(title = "Frequency of Reviews Over Months",
       x = "",
       y = "",
       fill = "Trip Type") +
  theme_minimal() +  theme(plot.title = element_text(hjust = 0.5), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```
We assumed that in summer there would be less business and more leisure trips because people take vacation in the season with better weather. In fact, surprisingly, in each month reviewers having leisure trip were significanlty more. Also, we see that in autumn there are significantly more business trips.





\newpage
We also assumed that if a person includes a picture in their review, it may either mean they liked the hotel so much and wanted to share beautiful pictures from there or they are disappointed because of something (maybe the rooms were dirty, or something was broken) and wanted to share that with a photo. We are checking which hypothesis is true usinq boxplots.
```{r datasetup5, echo=FALSE}
# Replace empty strings with NA in the 'images' column
group_5_hotels <- group_5_hotels %>%
  mutate(images = na_if(images, ""))

group_5_hotels <- group_5_hotels %>%
  mutate(has_image = ifelse(is.na(images), "With Image", "Without Image")) %>%
  mutate(has_image = factor(has_image))
```

```{r}
ggplot(group_5_hotels, aes(x = has_image, y = avg_rating)) +
  geom_boxplot() +
  labs(title = "Comparison of Reviews Containing Image by Average Rating",
       x = "",
       y = "") + theme_minimal() + theme() + theme(plot.title = element_text(hjust = 0.5))
```
People that icnluded images in their review gave higher ratings, according to the visualization. Let's also check the hypothesis statistically using linear regression:


```{r}
linear_model <- lm(avg_rating ~ has_image, data = group_5_hotels)
summary(linear_model)
```
P-value of "has_image" coefficient is very small meaning it's statistically significant and we retain our hypothesis.


Next thing we decided to do is to consider the length of the review texts. We assumed that if a person wrote a lot, it's possible they are complaining about some drawbacks, that's why their review is long. We stated a hypothesis that people who gave longer reviews rated hotels lower.
```{r datasetup6, echo=FALSE}
# Create a new column 'word_count' for reviews text
group_5_hotels <- group_5_hotels %>%
  mutate(word_count = str_count(review_text, "\\w+"))

# Calculate the IQR for word_count
iqr <- IQR(group_5_hotels$word_count)

# Define the lower and upper bounds to exclude outliers
lower_bound <- quantile(group_5_hotels$word_count)[2] - 1.5 * iqr
upper_bound <- quantile(group_5_hotels$word_count)[4] + 1.5 * iqr

# Filter the data to exclude outliers
group_5_hotels_filtered <- group_5_hotels %>%
  filter(word_count >= lower_bound & word_count <= upper_bound)

```




```{r}
# Create the plot with the filtered data
ggplot(group_5_hotels_filtered, aes(y = word_count, x = rating)) +
  geom_point() +
  labs(title = "Rating vs. Review Word Count",
       y = "Word Count",
       x = "Rating") +
  geom_smooth(se = FALSE) +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```
As we can see, indeed, reviews with lower ratings contain more words than reviews with higher ratings. The fitted lined shows that trend.

```{r}
linear_model <- lm(word_count ~ rating, data = group_5_hotels_filtered)
summary(linear_model)
```
Statistical analysis confirmed that there is such correlation: more words mean worse rating.


Time series analysis
```{r datasetup7, echo=FALSE}
# Convert the date to a Date object and create a year-month variable
group_5_hotels$reviewed_at <- as.Date(group_5_hotels$reviewed_at)
group_5_hotels$year_month <- floor_date(group_5_hotels$reviewed_at, "month")

# Group by year_month to get average rating per month
monthly_ratings <- group_5_hotels %>%
  group_by(year_month) %>%
  summarise(avg_rating = mean(rating, na.rm = TRUE))

# Create a complete sequence of months from start to end
all_months <- seq(min(monthly_ratings$year_month),
                  max(monthly_ratings$year_month), by = "month")

# Ensure all months are represented, even if no reviews were given
monthly_ratings_complete <- data.frame(year_month = all_months) %>%
  left_join(monthly_ratings, by = "year_month")

```

```{r}
# Plot the time series of average ratings per month
ggplot(monthly_ratings_complete, aes(x = year_month, y = avg_rating)) +
  geom_line(na.rm = TRUE) +  # Draw lines
  geom_point(na.rm = TRUE) +  # Add points
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +  
  # Set breaks to 1 month
  theme_minimal() +
  labs(title = "Monthly Average Ratings",
       x = "",
       y = "") +
  # Rotate x-axis labels for better readability
  theme(axis.text.x = element_text(angle = 90, hjust = 1), panel.grid.minor = element_blank(), )  
```
This shows the actual observed data, which seems to be fluctuating around a value just below 8.5 and doesn't show any clear long-term upward or downward trend within the displayed time frame.

```{r datasetup8, echo=FALSE}
# Convert to a ts object for STL decomposition
start_year <- year(min(monthly_ratings$year_month))
start_month <- month(min(monthly_ratings$year_month))
frequency <- 12 # Monthly data

monthly_ts <- ts(monthly_ratings$avg_rating, start=c(start_year, start_month), frequency=frequency)
```

```{r}
ggseasonplot(monthly_ts, year.labels = F) + 
  ggtitle("Seasonal Plot - Hotels Average Ratings") +
  xlab("") +
  ylab("") +
  theme(panel.grid.minor = element_blank(), 
        plot.title = element_text(size = 14, hjust = 0.5)) +  
  geom_line(linetype = "solid", size = 0.7) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5),
    axis.text.x = element_text(angle = 45, vjust = 1),
    legend.position = "bottom",
    legend.title = element_blank()
  ) 
```
We see that in 2018 and 2019 ratings were higher. In 2020, when COVID started, average ratings went lower, in May 2020 it was at its worst. It was the peak of the epidemic which may negatively affect people mood that's why they gave lower ratings and average ratings were that low. In 2021 the data is more stable excluding March. Overall, 5 years ago, average ratings were higher than then.

```{r datasetup9, echo=FALSE}
# Assuming the 'decomposed' variable contains the STL decomposition of your time series
decomposed <- stl(monthly_ts, s.window = 'periodic')

# Get the range of data without NA values for plotting
time_range <- time(decomposed$time.series)

# Convert July 2018 to the numeric time scale used by the time series object
start_time <- 2018 + (7 - 1) / 12  # July is the 7th month, so (7-1)/12

```

```{r}
# Plot the decomposed object using autoplot
autoplot(decomposed) +
  ggtitle("STL Decomposition of Hotels by Average Ratings") +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5),
    axis.title = element_blank(),
    axis.text = element_text(size = 10),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  xlim(c(start_time, max(time_range)))  # Set x-axis limits from July 2018 to the maximum of your time range
```


There's a noticeable downward trend starting from early 2020. This could reflect a variety of factors, including a possible impact of external events (such as the COVID-19 pandemic) on hotel ratings.

There's a repeating pattern that seems to occur on a regular basis throughout the period. This seasonality could correspond to factors like holidays, vacation seasons, or other cyclical events that influence hotel ratings.































